---
title: 'Laboratory Exercise Week 3'
author: "Ryan Estes | 06"
date: "09/06/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed laboratory exercise using Blackboard's Turnitin feature. Your Turnitin upload link is found on your Blackboard Course shell under the Laboratory folder.

***

For this exercise, you will need to use the package `mosaic` to find numerical and graphical summaries.

```{r}
# if mosaic package is not installed, then run install.packages("mosaic") 
# in your console: note that you need to this only one time
library(mosaic) # load the package to use its functions
```


1. Recall the `iris` data set from last week's exercise. The `iris` data set is already pre-loaded in R - look at the help file using `?iris` for more information on this data set.   
      i) Check the structure of the data using the function `str(iris)`.   
      ii) Find the average (or mean) measurement of the variable `Sepal.Length`. Do this in two ways as described in the lesson.   
      iii) Find the average `Sepal.Length` for the different flower `Species`. Give a brief comment on the averages.  
      iv) Repeat (ii) and (iii) but use the summary standard deviation `sd()` which describes the spread of the variable.   
      v) Describe the shape of the variable `Sepal.Length` by creating a histogram using `histogram()`.  Write your description outside the code chunk.  
      vi) Compare the `Sepal.Length` of the three species of flowers by creating a side-by-side boxplot using `bwplot()`. Write your description outside the code chunk.
      

### Code chunk
```{r} 
# Insert your code for this question after this line
# i)
str(iris)

# ii)
mean(iris$Sepal.Length)
#mean(~ Sepal.Length, data = iris)

# iii)
# iv)
# v)
library(plyr)
ddply(iris, ~Species,summarise,mean=mean(Sepal.Length),sd=sd(Sepal.Length))
#mean(Species ~ Sepal.Length, data = iris)

# Mean value of Sepal.Length of Setosa < Versicolor < Virginica

# iv) 
sd(iris$Sepal.Length)
#sd(~ Sepal.Length, data = iris)

# v)
hist(iris$Sepal.Length)

# vi)
boxplot(Sepal.Length ~ Species, data = iris)

# last R code line
```

     Species  mean        sd
1     setosa 5.006 0.3524897
2 versicolor 5.936 0.5161711
3  virginica 6.588 0.6358796

v) The histogram looks similar to a bell curve with most of the Lengths between 4.5-7. The most Lengths lie between 6-6.5.

vi) The Boxplot description:

Setosa:
  Min: 4.2
  1st Quartile: 4.8
  Median: 5.0
  3rd Quartile: 5.2
  Max: 5.8
  
Versicolor:
  Min: 4.9
  1st Quartile: 5.7
  Median: 6.0
  3rd Quartile: 6.4
  Max: 7.1

Virginica:
  Min: 5.7
  1st Quartile: 6.3
  Median: 6.6
  3rd Quartile: 7.0
  Max: 7.9


2. The data set `MLB-TeamBatting-S16.csv` contains MLB Team Batting Data for selected variables. Load the data set from the given url using the code below. This data set was obtained from [Baseball Reference](https://www.baseball-reference.com/leagues/MLB/2016-standard-batting.shtml).
    * Tm - Team   
    * Lg - League: American League (AL), National League (NL)  
    * BatAge - Batters’ average age  
    * RPG - Runs Scored Per Game  
    * G - Games Played or Pitched  
    * AB - At Bats  
    * R - Runs Scored/Allowed  
    * H - Hits/Hits Allowed  
    * HR - Home Runs Hit/Allowed  
    * RBI - Runs Batted In  
    * SO - Strikeouts  
    * BA - Hits/At Bats  
    * SH - Sacrifice Hits (Sacrifice Bunts)  
    * SF - Sacrifice Flies  

    i) Find the average measurement for the following variables `BatAge`, `RPG`, `R`, `H` and `BA`.   
    ii) Create dotplot's or histogram's for each variable in (i).     
    iii) Using your own words, describe the distribution of each variable in (i).   Write your answer outside the code chunk.     
    iv) Find the average and the standard deviation of the variables `RPG`, `H` and `BA` for each league.     
    v) Describe any differences or similarities between the leagues. Write your comment outside the code chunk.
      


### Code chunk
```{r} 
# load the data set
mlb16.data <- read.csv("https://raw.githubusercontent.com/jpailden/rstatlab/master/data/MLB-TeamBatting-S16.csv")
str(mlb16.data) # check structure
head(mlb16.data)  # show first six rows

# i)
mean(mlb16.data$BatAge)
mean(mlb16.data$RPG)
mean(mlb16.data$R)
mean(mlb16.data$H)
mean(mlb16.data$BA)

# ii)
hist(mlb16.data$BatAge)
hist(mlb16.data$RPG)
hist(mlb16.data$R)
hist(mlb16.data$H)
hist(mlb16.data$BA)

# iii)
# Below

# iv)

library(plyr)
ddply(mlb16.data, ~Lg,summarise,mean=mean(BatAge),sd=sd(BatAge))
ddply(mlb16.data, ~Lg,summarise,mean=mean(RPG),sd=sd(RPG))
ddply(mlb16.data, ~Lg,summarise,mean=mean(R),sd=sd(R))
ddply(mlb16.data, ~Lg,summarise,mean=mean(H),sd=sd(H))
ddply(mlb16.data, ~Lg,summarise,mean=mean(BA),sd=sd(BA))

# last R code line
```

  #ddply(mlb16.data, ~Lg,summarise,mean=mean(BatAge),sd=sd(BatAge))
  
  Lg     mean        sd
1 AL 28.64667 1.0676186
2 NL 28.21333 0.8458864

The AL had a higher age but skewed more based on standard deviation.

  #ddply(mlb16.data, ~Lg,summarise,mean=mean(RPG),sd=sd(RPG))
  
  Lg     mean        sd
1 AL 4.519333 0.3533607
2 NL 4.437333 0.3914345

THE AL had, on average, more runs per game than the NL as well as a lower sd.

  #ddply(mlb16.data, ~Lg,summarise,mean=mean(R),sd=sd(R))
  
  Lg     mean       sd
1 AL 731.3333 56.83016
2 NL 718.2667 63.89105

The AL had more runs per game than the NL as well as a lower sd.

  #ddply(mlb16.data, ~Lg,summarise,mean=mean(H),sd=sd(H))
  
  Lg     mean       sd
1 AL 1419.933 64.49858
2 NL 1398.467 71.57301

The AL had more hits per than the NL as well as a lower sd.

  #ddply(mlb16.data, ~Lg,summarise,mean=mean(BA),sd=sd(BA))
  
  Lg      mean          sd
1 AL 0.2568667 0.009869626
2 NL 0.2537333 0.009837731

The AL had a slightly better batting average than the NL, but they are basically the same. .257 vs .254

iii)

  #mean(mlb16.data$BatAge)
  
[1] 28.43

The average batter's age in the MLB for 2016 was about 28 and a half years old.

  #mean(mlb16.data$RPG)
  
[1] 4.478333

The average runs per game in the MLB for 2016 was a little under 5.

  #mean(mlb16.data$R)
  
[1] 724.8

The average runs scored per team in the MLB for 2016 was about 725 for the season.

  #mean(mlb16.data$H)
  
[1] 1409.2

The average hits per team in the MLB for 2016 was about 1409 for the season.

  #mean(mlb16.data$BA)
  
[1] 0.2553

The average batting average per team in the MLB for 2016 was about .255 (or a hit 26% of the time) for the season.
