---
title: 'Laboratory Exercise Week 2'
author: "Ryan Estes | Section 006"
date: "8/30/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed laboratory exercise using Blackboard's Turnitin feature. Your Turnitin upload link is found on your Blackboard Course shell under the Laboratory folder.

***

1. You will work with a matrix whose entries are all prime numbers below 30.
      i)  Create this matrix using the function `matrix()` with five rows. Save the matrix as `P`.
      ii) Extract the second and third row out of `P`.
      iii) Extract the entry in the fourth row and first column of `P.
      iv) Generate the transpose of the matrix using the function `t()`. What is the new dimension of this matrix.

## Code chunk
```{r}
# Insert your code for this question after this line
#i
P <- matrix(c(2,3,5,7,11,13,17,19,23,29), nrow = 5)
 
#ii
row2 <- P[2, ]
row3 <- P[3, ]

#iii
row4col1 <- P[4,1]

#iv
transpose <- t(P)
## The new dimension is a 5 column matrix instread of 5 rows. 
```

2. Create and manipulate a data frame.
      i) Create a `data.frame` named `my.trees` that has the following columns:
          - `Girth = c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0)`
          - `Height = c(70, 65, 63, 72, 81, 83, 66)`
          - `Volume = c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)`
      ii) Extract the fifth observational unit.
      iii) Extract the `Girth` column referring to it by name.
      iv) Print out a data frame of all the observations except for the last observation.


## Code chunk
```{r}
# Insert your code for this question after this line
#i
my.trees <- data.frame(
  Girth = c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0),
  Height = c(70, 65, 63, 72, 81, 83, 66),
  Volume = c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)
)

#ii
my.trees[5,]

#iii
my.trees[, "Girth"]

#iv
print(my.trees[1:6,])
```

3. The `RailTrail` dataset within the `mosaic` package includes the counts of crossings of a rail trail in Northampton, Massachusetts for 90 days in 2005. City officials are interested in understanding usage of the trail network, and how it changes as a function of temperature and day of the week.
      i) Check the structure of the`RailTrail`. 
      ii) How many variables and observations are in the data set? 
      iii) Which variables are `integer` type?
      iv) Display the first 4 rows of the data.

## Code chunk
```{r} 
# Insert your code for this question after this line
#install.packages("mosaic") #needed for RailTrail

#i
RailTrail <- mosaicData::RailTrail
str(RailTrail)

#ii
#RailTrail contains 90 observations and 10 variables.

#iii
#Variables of type integer:
#hightemp
#lowtemp
#spring
#summer
#fall
#volume

#iv
head(RailTrail, 4)

# last R code line
```



  
4.  You will work with the passenger Titanic data found on this [link](<https://goo.gl/NHb1Pg>). 
      i) Read this data directly from the web using its URL.
      ii) Read this data by saving it first into your working directory. Check your working directory using `getwd()` or change its location using the steps covered in the lesson.
      iii) How many passengers and what variables are in the data set? 
      iv) Display the first 5 rows of the data.

## Code chunk
```{r}
# Insert your code for this question after this line
#i
TitanicPass.data <- read.csv("https://goo.gl/NHb1Pg", header = TRUE)

#ii
currentdir <- getwd()
LocalTitanicPass.data <- "/localtitanicdata.csv"
write.table(TitanicPass.data, file=paste(currentdir, LocalTitanicPass.data, sep = ""))

#iii
#passengers = 1316
#variables include:
#X
#class
#age
#sex
#survived

#iv
head(TitanicPass.data, 5)
```

